---
title: "PGY1 - Allocations 2024"
author: "Dr Scott Johnson Gelzinnis"
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: journal
    df_print: paged
  html_document:
    theme: journal
    df_print: paged
  pdf_document: default
---

*Copyright Dr Scott Johnson Gelzinnis (2023) - Statement below*


  
# Setting up the Environment  


```{r include=TRUE, echo=TRUE, results='hide'}

library(here())
i_am("PGY2_Solve_TRIAL.Rmd")

# List of required packages
required_packages <- c("randomNames", "lpSolve", "lpSolveAPI", "ROI", "gt", "readr", "tidyverse", "googlesheets4", "stringr", "knitr", "writexl")

# Check if packages are installed and install them if they are not
new_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

# Load the packages
lapply(required_packages, require, character.only = TRUE)

# Random seed
set.seed(69)

```



# Loading in the Preference Matrix 

```{r include=TRUE, echo=TRUE, results='hide'}

## Read in Google Sheet
raw_preference_data <- read_sheet("https://docs.google.com/spreadsheets/d/19xMulDMIK6FgmfMajmI1BnlaSsVut_idzsIN1k-X-Aw/edit?usp=sharing")

# Convert the tibble to a data.frame
raw_preference_data <- as.data.frame(raw_preference_data)

## Define preference matrix  ##

# Combine columns 13 and 14 to create the "Doctor" column
raw_preference_data$Doctor <- paste(raw_preference_data[[13]], raw_preference_data[[14]], sep = ", ")

# Sort the raw_preference_data by Doctor column
raw_preference_data <- raw_preference_data %>%
  arrange(Doctor)

# Extract column names that have the pattern '| '
term_columns <- grep("\\| ", names(raw_preference_data), value = TRUE)

# Select the relevant columns
selected_data <- raw_preference_data[, term_columns]

# Select survey data and load as a data frame
# Read in and rename columns of survey_data
survey_data <- raw_preference_data %>%
  select(
    Doctor, 
    `Desired Career Pathway (GP, Medicine, Surgery, Critical Care, Psych)`, 
    `Do you have plans requiring leave of more than 2 weeks in the 2024 clinical year? (If possible your request will be accommodated)`
  ) %>%
  rename(
    `Desired Career Pathway` = `Desired Career Pathway (GP, Medicine, Surgery, Critical Care, Psych)`,
    `Leave Request Term` = `Do you have plans requiring leave of more than 2 weeks in the 2024 clinical year? (If possible your request will be accommodated)`
  )

# Set "Doctor" column as row names
rownames(selected_data) <- raw_preference_data$Doctor

# Clean the column names by removing everything before '| ' (i.e., pipe followed by a space)
colnames(selected_data) <- gsub(".*\\| ", "", colnames(selected_data))

# Convert selected data to a matrix and exclude the "Doctor" column
preference_matrix <- as.matrix(selected_data)

# Define dimensions based on preference_matrix
num_columns <- ncol(preference_matrix)
num_doctors <- nrow(preference_matrix)
num_terms <- num_columns  

# Assuming the number of terms is equal to the number of columns in the matrix
row_names <- rownames(selected_data)
column_names <- colnames(selected_data)

```


# Definitions

```{r include=TRUE, echo=TRUE, results='hide'}
               
                         ## Linear Optimization - Defining the problem ##

term_master_df <- read_csv("PGY1 2024 Dataframe.csv")

# Directly infer the number of doctors and terms from the preference matrix
num_doctors <- nrow(preference_matrix)
num_terms <- ncol(preference_matrix)

# Define the objective function coefficients (preferences)
objective_coeffs <- as.vector(preference_matrix)

# Extract the maximum number of doctors allowed for each term using the column names
max_doctors_per_term1 <- as.vector(term_master_df$`PGY1 Term 1 Allocation`)
max_doctors_per_term2 <- as.vector(term_master_df$`PGY1 Term 2 Allocation`)
max_doctors_per_term3 <- as.vector(term_master_df$`PGY1 Term 3 Allocation`)
max_doctors_per_term4 <- as.vector(term_master_df$`PGY1 Term 4 Allocation`)
max_doctors_per_term5 <- as.vector(term_master_df$`PGY1 Term 5 Allocation`)

# Helper function for tidy conversions
convert_and_name <- function(column_name) {
  term_master_df %>%
    pull(column_name) %>%
    as.factor() %>%
    set_names(term_master_df$`Term Name`)
}

# Using the helper function to define the vectors
clinical_structure_term <- convert_and_name('Clinical Team Structure')
specialty_status_per_term <- convert_and_name('Specialty Status')
sub_specialty_status_per_term <- convert_and_name('Subspecialty Status')
rural_status_term <- convert_and_name('Rurality')
term_classification1 <- convert_and_name('Term Classification')
term_classification2 <- convert_and_name('Term Classification 2')

```



# Linear Optimization Model with Minimized Dissatisfaction

### With allowable constraints to provide viable solution (i.e if unable to solve, roll back constraint one at a time)

```{r include=TRUE, echo=TRUE, results='hide'}


## Linear Optimization with Weighted Dissatisfaction Values ##

# Use the preference_matrix directly as the dissatisfaction_values
dissatisfaction_values <- preference_matrix

# Convert dissatisfaction values to a vector for the objective function
objective_coeffs_dissatisfaction <- as.vector(dissatisfaction_values)

# Extend LP model to include slack variables
lp_model_dissatisfaction <- make.lp(0, num_doctors * num_terms + num_doctors)

# SLACK VARIABLE TRIAL
slack_penalty <- max(dissatisfaction_values) * 10
slack_coefficients <- rep(slack_penalty, num_doctors)
objective_coeffs_extended <- c(objective_coeffs_dissatisfaction, slack_coefficients)

# Set binary decision variables (0 or 1)
set.type(lp_model_dissatisfaction, columns = 1:(num_doctors * num_terms), type = "binary")

# Set the objective function coefficients for minimizing dissatisfaction
set.objfn(lp_model_dissatisfaction, objective_coeffs_extended)



## NEXT STEP TO RUN A LOOP TO ASSIGN 5 TERMS ACROSS THE YEAR BASED ON PREFERENCES ##

# Number of times to run the assignment
num_assignments <- 5

# Initialize a list to store the results of each assignment
assignment_results <- list()

# Create a list to store unassigned terms for each assignment iteration
unassigned_terms_list <- list()

# Matrix to track number of specialty terms assigned to each doctor
# Get unique specialty statuses from the 'specialty_status_per_term' vector
unique_specialty_statuses <- unique(specialty_status_per_term)
# Rows correspond to doctors, columns correspond to specialties
specialty_assignments_tracker <- matrix(0, nrow = num_doctors, ncol = length(unique_specialty_statuses))
# The columns should correspond to the unique specialties
colnames(specialty_assignments_tracker) <- unique_specialty_statuses


# This matrix will have rows corresponding to doctors and columns corresponding to terms.
# An entry of 1 in the matrix indicates that the doctor (row) has been assigned the term (column).
assigned_terms_matrix <- matrix(0, nrow = num_doctors, ncol = num_terms)

# Loop through the assignments
for (assignment_num in 1:num_assignments) {
  # Create an empty LP model for minimizing dissatisfaction
lp_model_dissatisfaction <- make.lp(0, num_doctors * num_terms + num_doctors)
  
  # Set binary decision variables (0 or 1)
  set.type(lp_model_dissatisfaction, columns = 1:(num_doctors * num_terms), type = "binary")
  
  # Set the objective function coefficients for minimizing dissatisfaction
  set.objfn(lp_model_dissatisfaction, objective_coeffs_extended)
  
  # Add constraints to ensure one assignment per doctor
for (i in 1:num_doctors) {
    constr_one_assignment <- rep(0, num_doctors * num_terms + num_doctors)
    constr_one_assignment[((i - 1) * num_terms + 1):(i * num_terms)] <- 1
    add.constraint(lp_model_dissatisfaction, constr_one_assignment, type = "=", rhs = 1)
}

  # Add constraint for top 10 terms for each doctor (MODIFIED TO ADD SLACK)
margin = -5
for (i in 1:num_doctors) {
    top_10_terms_for_doctor <- order(preference_matrix[i,])[1:(10+margin)]
    constr_top_10 <- rep(0, num_doctors * num_terms + num_doctors)
    for (j in top_10_terms_for_doctor) {
        constr_top_10[(i - 1) * num_terms + j] <- 1
    }
    constr_top_10[num_doctors * num_terms + i] <- -margin
    add.constraint(lp_model_dissatisfaction, constr_top_10, type = ">=", rhs = 1)
}
    
# 1. Max doctors per term constraint:

# Helper function to retrieve the correct 'max doctors' vector for a given term
   get_max_doctors_for_term <- function(term) {
     switch(term,
            "1" = max_doctors_per_term1,
            "2" = max_doctors_per_term2,
            "3" = max_doctors_per_term3,
            "4" = max_doctors_per_term4,
            "5" = max_doctors_per_term5)
   }


for (j in 1:num_terms) {
  constr_max_doctors_per_term <- rep(0, num_doctors * num_terms + num_doctors)
  
  indices <- j + seq(0, (num_doctors - 1) * num_terms, by = num_terms)
  constr_max_doctors_per_term[indices] <- 1
  
  max_doctors_for_this_term <- get_max_doctors_for_term(assignment_num)
  add.constraint(lp_model_dissatisfaction, constr_max_doctors_per_term, type = "<=", rhs = max_doctors_for_this_term[j])
}

# 2. Constraint to prevent reassignment to same term:
for (i in 1:num_doctors) {
  for (j in 1:num_terms) {
    if (assigned_terms_matrix[i, j] == 1) {
      constraint_already_assigned <- rep(0, num_doctors * num_terms + num_doctors)
      constraint_already_assigned[(i - 1) * num_terms + j] <- 1
      add.constraint(lp_model_dissatisfaction, constraint_already_assigned, type = "=", rhs = 0)
    }
  }
}

# 3. Specialty Status Constraint

strict_limit <- 2
   
for (specialty in unique_specialty_statuses) {
  for (i in 1:num_doctors) {
    constr_specialty <- rep(0, num_doctors * num_terms + num_doctors)
    specialty_indices <- which(specialty_status_per_term == specialty)
    
    # Assign 1 to the constraint vector for indices related to the current specialty
    constr_specialty[(i - 1) * num_terms + specialty_indices] <- 1
    
    # Calculate the remaining number of assignments allowed for this specialty for the doctor
    # Ensure that a doctor does not get assigned more than a strict limit from the same specialty across all iterations
    terms_already_assigned <- specialty_assignments_tracker[i, specialty]
    remaining_assignments_allowed <- max(0, strict_limit - terms_already_assigned)
    
    # Update the right-hand side of the constraint for the current iteration
    add.constraint(lp_model_dissatisfaction, constr_specialty, type = "<=", rhs = remaining_assignments_allowed)
  }
}


# 
# # 4. Sub-Specialty Status Constraint:
# unique_sub_specialties <- unique(sub_specialty_status_per_term[!is.na(sub_specialty_status_per_term)])
# for (sub_specialty in unique_sub_specialties) {
#   for (i in 1:num_doctors) {
#     constr_sub_specialty <- rep(0, num_doctors * num_terms + num_doctors)
#     for (j in which(sub_specialty_status_per_term == sub_specialty)) {
#       constr_sub_specialty[(i - 1) * num_terms + j] <- 1
#     }
#     add.constraint(lp_model_dissatisfaction, constr_sub_specialty, type = "<=", rhs = 1)
#   }
# }
  
#  # Constraint for >50% in team based 
  
  
  # Solve the linear programming problem to minimize dissatisfaction
  lp_result_dissatisfaction <- solve(lp_model_dissatisfaction)
  
  # Check if the LP problem was solved successfully
  if (lp_result_dissatisfaction == 0) {
  optimal_solution_dissatisfaction <- get.variables(lp_model_dissatisfaction)
  assignment_results[[assignment_num]] <- optimal_solution_dissatisfaction

  # Update the assigned_terms_matrix based on this solution
  for (i in 1:num_doctors) {
    for (j in 1:num_terms) {
      if (optimal_solution_dissatisfaction[(i - 1) * num_terms + j] == 1) {
        assigned_terms_matrix[i, j] <- 1
        # Assuming that you have a vector/matrix that links terms to specialties
        term_specialty <- specialty_status_per_term[j]  # specialty_status_per_term needs to be defined earlier in your code

        # Assuming that your specialty_assignments_tracker has row names as doctors and column names as specialties
        # Here we increment the count for the specialty corresponding to the term assigned in this iteration
        specialty_assignments_tracker[i, term_specialty] <- specialty_assignments_tracker[i, term_specialty] + 1
      }
    }
  }
} else {
  cat("LP problem could not be solved for assignment", assignment_num, "Result:", lp_result_dissatisfaction, "\n")
 }
}

# Print the results of each assignment
for (assignment_num in 1:num_assignments) {
  cat("Assignment", assignment_num, "results:\n")
  optimal_solution_dissatisfaction <- assignment_results[[assignment_num]]
  for (i in 1:num_doctors) {
    for (j in 1:num_terms) {
      if (optimal_solution_dissatisfaction[(i - 1) * num_terms + j] == 1) {
        doctor_name <- row_names[i]
        term_name <- column_names[j]
        cat("Doctor", doctor_name, "is assigned to Term", term_name, "\n")
      }
    }
  }
  cat("\n")
}
# Initialize a data frame to store the results
assignment_table <- data.frame(Doctor = row_names, 
                               Assignment1 = character(num_doctors), 
                               Assignment2 = character(num_doctors),
                               Assignment3 = character(num_doctors),
                               Assignment4 = character(num_doctors),
                               Assignment5 = character(num_doctors),
                               stringsAsFactors = FALSE)

# Populate the data frame with the assignment results
for (assignment_num in 1:num_assignments) {
  optimal_solution_dissatisfaction <- assignment_results[[assignment_num]]
  for (i in 1:num_doctors) {
    for (j in 1:num_terms) {
      if (optimal_solution_dissatisfaction[(i - 1) * num_terms + j] == 1) {
        term_name <- column_names[j]
        assignment_column <- paste0("Assignment", assignment_num)
        assignment_table[i, assignment_column] <- term_name
      }
    }
  }
}


# Print the table using gt
gt(assignment_table)


```



# PGY2 Unassigned terms Data Frame and Terms List
```{r}

# Initialize a data frame to store the unassigned slots for each term
unassigned_terms_table <- data.frame(Term = column_names, 
                                     Assignment1 = integer(length(column_names)), 
                                     Assignment2 = integer(length(column_names)),
                                     Assignment3 = integer(length(column_names)),
                                     Assignment4 = integer(length(column_names)),
                                     Assignment5 = integer(length(column_names)),
                                     stringsAsFactors = FALSE)

# Helper function to get total assignments for a term for a given assignment number
get_total_assignments_for_term <- function(term_index, assignment_num) {
  optimal_solution_dissatisfaction <- assignment_results[[assignment_num]]
  return(sum(optimal_solution_dissatisfaction[seq(term_index, num_doctors * num_terms, by = num_terms)]))
}

# Compute the unassigned slots for each term and assignment
for (term_index in 1:length(column_names)) {
  for (assignment_num in 1:num_assignments) {
    max_doctors_for_this_term <- get_max_doctors_for_term(assignment_num)
    total_assignments <- get_total_assignments_for_term(term_index, assignment_num)
    
    unassigned_slots <- max_doctors_for_this_term[term_index] - total_assignments
    
    assignment_column <- paste0("Assignment", assignment_num)
    unassigned_terms_table[term_index, assignment_column] <- unassigned_slots
  }
}

# View the unassigned_terms_table
gt(unassigned_terms_table)

                                            # Generate this data frame as a term list

# Initialize a new dataframe for unassigned terms
# Drop the first column for calculation purposes
numeric_unassigned_terms <- unassigned_terms_table[,-1]

# Compute the maximum unassigned terms across all assignments
max_unassigned <- max(as.matrix(numeric_unassigned_terms))

# Create a list to store term repetitions for each assignment
unassigned_list <- list()

for (assignment_num in 1:num_assignments) {
  assignment_column <- paste0("Assignment", assignment_num)
  terms_rep <- rep(unassigned_terms_table$Term, numeric_unassigned_terms[[assignment_column]])
  unassigned_list[[assignment_num]] <- terms_rep
}

# Make all lists the same length by padding with NA
max_length <- max(sapply(unassigned_list, length))
unassigned_list <- lapply(unassigned_list, function(x) {
  c(x, rep(NA, max_length - length(x)))
})

# Convert the list to dataframe
unassigned_df <- as.data.frame(unassigned_list)
names(unassigned_df) <- paste0("Term", 1:num_assignments)

# View the dataframe
gt(unassigned_df)


```


# MODEL EVALUATION

```{r include=TRUE, echo=TRUE, results='markup'}

# How many Doctors with one Top 10 preference

# Convert the preference matrix into a top 10 boolean matrix
top_10_matrix <- t(apply(preference_matrix, 1, function(row) row %in% sort(row, decreasing = FALSE)[1:10]))

# For each doctor, determine if the assigned terms are in their top 10 preferences
assignment_top_10_matrix <- matrix(0, nrow = num_doctors, ncol = num_assignments)

for (assignment_num in 1:num_assignments) {
  assignment_column <- paste0("Assignment", assignment_num)
  assigned_terms <- assignment_table[, assignment_column]
  
  # Convert the assigned term names back to their column indices
  assigned_indices <- match(assigned_terms, colnames(selected_data))
  
  for (i in 1:num_doctors) {
    assignment_top_10_matrix[i, assignment_num] <- top_10_matrix[i, assigned_indices[i]]
  }
}

# Convert the preference matrix into a top 5 boolean matrix
top_5_matrix <- t(apply(preference_matrix, 1, function(row) row %in% sort(row, decreasing = FALSE)[1:5]))

# Create a function to compute metrics for a given top_x matrix
get_metrics <- function(matrix, assignment_table, num_assignments) {
  assignment_matrix <- matrix(0, nrow = num_doctors, ncol = num_assignments)
  
  for (assignment_num in 1:num_assignments) {
    assignment_column <- paste0("Assignment", assignment_num)
    assigned_terms <- assignment_table[, assignment_column]
    
    # Convert the assigned term names back to their column indices
    assigned_indices <- match(assigned_terms, colnames(selected_data))
    
    for (i in 1:num_doctors) {
      assignment_matrix[i, assignment_num] <- matrix[i, assigned_indices[i]]
    }
  }
  
  return(rowSums(assignment_matrix))
}

top_10_counts_per_doctor <- get_metrics(top_10_matrix, assignment_table, num_assignments)
top_5_counts_per_doctor <- get_metrics(top_5_matrix, assignment_table, num_assignments)

# Extract all assigned terms for each doctor across the assignments
assigned_terms_all <- as.vector(t(assignment_table[, 2:(num_assignments + 1)]))

# Convert the assigned term names back to their column indices
assigned_indices_all <- match(assigned_terms_all, colnames(selected_data))

# Retrieve preference values using matrix indexing
assigned_preference_values <- matrix(0, nrow = num_doctors, ncol = num_assignments)
for (assignment_num in 1:num_assignments) {
  assignment_column <- paste0("Assignment", assignment_num)
  assigned_terms <- assignment_table[, assignment_column]
  
  # Convert the assigned term names back to their column indices
  assigned_indices <- match(assigned_terms, colnames(selected_data))
  
  for (i in 1:num_doctors) {
    assigned_preference_values[i, assignment_num] <- preference_matrix[i, assigned_indices[i]]
  }
}

average_assigned_preference <- mean(as.vector(assigned_preference_values))

# Calculate the metrics of interest for each doctor
range_per_doctor <- apply(assigned_preference_values, 1, range)
max_per_doctor <- range_per_doctor[2,]
min_per_doctor <- range_per_doctor[1,]
sd_per_doctor <- apply(assigned_preference_values, 1, sd)

# Compute the average values across all doctors
average_max = mean(max_per_doctor)
average_min = mean(min_per_doctor)
average_sd = mean(sd_per_doctor)

# Create a data frame with the evaluation metrics
evaluation_data <- data.frame(
  Metric = c("At least 1 term in Top 10 (%)", 
             "At least 2 terms in Top 10 (%)", 
             "At least 3 terms in Top 10 (%)", 
             "At least 1 term in Top 5 (%)",
             "Average Preference of Assigned Terms",
             "Average Max of Assigned Preferences",
             "Average Min of Assigned Preferences",
             "Average SD of Assigned Preferences"),
  Value = c(sum(top_10_counts_per_doctor > 0) / num_doctors * 100, 
            sum(top_10_counts_per_doctor >= 2) / num_doctors * 100, 
            sum(top_10_counts_per_doctor >= 3) / num_doctors * 100, 
            sum(top_5_counts_per_doctor > 0) / num_doctors * 100,
            average_assigned_preference,
            average_max,
            average_min,
            average_sd)
)

# Round the values to 2 decimal places
evaluation_data$Value <- round(evaluation_data$Value, 2)

# Generate the table using gt
evaluation_table <- gt(evaluation_data) %>% 
  tab_header(
    title = "Model Evaluation"
  ) %>% 
  cols_label(
    Metric = "Metric",
    Value = "Value"
  )

# Print the table
print(evaluation_table)


```


                         
# Reporting - 

### To add transparancy to allocations process and to assist in manual auditing of lines in order to preserve the integrity of the preferences of the cohort. Assist with manual changes to make terms HETI compliant or understand GAPS in the event their is nil viable solution to whole gamut of constraints. 

```{r include=TRUE, echo=TRUE, results='markup'}
# 1. Doctors Preference Report

# Initialize an empty data frame to store the report
doctor_preference_report <- data.frame(
  Doctor = character(),
  Term = character(),
  Preference_Score = numeric(),
  stringsAsFactors = FALSE
)

# Iterate through each doctor
for (i in 1:num_doctors) {
  doctor_name <- row_names[i]
  
  # Iterate through each of the assignments to extract the allocated terms for each doctor
  for (assignment_num in 1:num_assignments) {
    optimal_solution_dissatisfaction <- assignment_results[[assignment_num]]
    for (j in 1:num_terms) {
      if (optimal_solution_dissatisfaction[(i - 1) * num_terms + j] == 1) {
        assigned_term <- column_names[j]
        
        # Access the doctor's preference row from the preference matrix
        doctor_preference <- preference_matrix[i, ]
        
        # Calculate preference score for the assigned term
        preference_score <- doctor_preference[match(assigned_term, column_names)]
        
        # Build a data frame with the results for this doctor and term
        doctor_report <- data.frame(
          Doctor = rep(doctor_name, 1),
          Term = assigned_term,
          Preference_Score = preference_score
        )
        
        # Append the doctor's report to the overall report data frame
        doctor_preference_report <- rbind(doctor_preference_report, doctor_report)
      }
    }
  }
}


gt(doctor_preference_report)
```


```{r include=TRUE, echo=TRUE, results='markup'}
# 2. HETI Compliance Report


# Calculate average preference for each doctor across all 5 assignments
avg_preference <- doctor_preference_report %>%
  group_by(Doctor) %>%
  summarize(Avg_Score = mean(Preference_Score, na.rm = TRUE))

# Check if a doctor has been assigned to more than 1 term in any sub-specialty
sub_specialty_compliance <- apply(assignment_table[,-1], 1, function(assignments) {
  # Map the assignments to their corresponding sub_specialty_status_per_term
  sub_specialty_terms <- sub_specialty_status_per_term[assignments]
  
  # Check if any sub-specialty is assigned more than 1 time
  max_count <- max(table(sub_specialty_terms), na.rm = TRUE)
  return(max_count <= 1)
})

# Check if a doctor has been assigned to more than 2 terms in any specialty
specialty_compliance <- apply(assignment_table[,-1], 1, function(assignments) {
  # Map the assignments to their corresponding specialty_status_per_term
  specialty_terms <- specialty_status_per_term[assignments]
  
  # Check if any specialty is assigned more than 2 times
  max_count <- max(table(specialty_terms), na.rm = TRUE)
  return(max_count < 3)
})

# Check if a doctor has been assigned to at least 3 "Team Based" terms
team_based_compliance <- sapply(1:num_doctors, function(i) {
  doctor_name <- row_names[i]
  terms_assigned <- doctor_preference_report$Term[doctor_preference_report$Doctor == doctor_name]
  
  # Debugging print statements
  print(paste("Doctor:", doctor_name))
  print(paste("Assigned terms:", toString(terms_assigned)))
  structure_indices <- match(terms_assigned, names(clinical_structure_term))
  print(paste("Matching indices:", toString(structure_indices)))
  print(paste("Clinical structures:", toString(clinical_structure_term[structure_indices])))
  
  team_based_count <- sum(clinical_structure_term[structure_indices] == "Team Based")
  
  return(team_based_count >= 3)
})

# Check for compliance with the desired term classifications
check_term_classifications <- function(terms_assigned) {
  classification1 <- term_classification1[match(terms_assigned, column_names)]
  classification2 <- term_classification2[match(terms_assigned, column_names)]
  required_terms <- c("A-Undifferentiated illness patient care", 
                      "B-Chronic illness patient care", 
                      "C-Acute critical illness patient care",
                      "D-Peri-operative / procedural patient care")
  
  # Check for each term individually, without merging the classifications
  has_required_terms <- sapply(required_terms, function(term) {
    term %in% classification1 || term %in% classification2
  })
  
  return(all(has_required_terms))
}

term_classifications_compliance <- sapply(1:num_doctors, function(i) {
  doctor_name <- row_names[i]
  terms_assigned <- doctor_preference_report$Term[doctor_preference_report$Doctor == doctor_name]
  return(check_term_classifications(terms_assigned))
})

compliance_data <- data.frame(
  Doctor = avg_preference$Doctor,
  Avg_Preference = avg_preference$Avg_Score,
  SubSpecialty_Compliance = sub_specialty_compliance,
  Specialty_Compliance = specialty_compliance,
  Team_Based_Compliance = team_based_compliance,
  Term_Classifications_Compliance = term_classifications_compliance
)

# Format with gt

HETI_compliance_table <- compliance_data %>%
  gt() %>%
  cols_label(
    Avg_Preference = "Avg. Preference (1-5)",
    SubSpecialty_Compliance = "Max 1 Term in Sub-Specialty",
    Specialty_Compliance = "Max 2 Terms in Specialty",
    Team_Based_Compliance = "At Least 3 'Team Based' Terms",
    Term_Classifications_Compliance = "Essential Classifications"
  )

print(HETI_compliance_table)
```

```{r}
export_tables <- function(assignment_table, evaluation_table, doctor_preference_report, HETI_compliance_table, preference_matrix, unassigned_df, filename) {
  
  # Convert gt tables back to data.frame if necessary
  if (inherits(assignment_table, "gt_tbl")) {
    assignment_table <- as.data.frame(assignment_table)
  }
  if (inherits(evaluation_table, "gt_tbl")) {
    evaluation_table <- as.data.frame(evaluation_table)
  }
  if (inherits(doctor_preference_report, "gt_tbl")) {
    doctor_preference_report <- as.data.frame(doctor_preference_report)
  }
  if (inherits(HETI_compliance_table, "gt_tbl")) {
    HETI_compliance_table <- as.data.frame(HETI_compliance_table)
  }
  
  # Convert the preference_matrix to a data frame and add Doctor names if it's a matrix
  if (is.matrix(preference_matrix)) {
    preference_matrix_df <- as.data.frame(preference_matrix)
    # Add doctor names from the rownames of the matrix if they exist
    if (!is.null(rownames(preference_matrix))) {
      preference_matrix_df <- cbind(Doctor = rownames(preference_matrix), preference_matrix_df)
    }
  } else {
    preference_matrix_df <- preference_matrix
  }
  
  # Join the survey_data and assignment_table on the "Doctor" column
  assignment_table_with_survey <- left_join(survey_data, assignment_table, by = "Doctor")
  
  # Prepare the list of tables to be exported
  tables_list <- list(
    AssignmentTable = assignment_table_with_survey,
    EvaluationTable = evaluation_table,
    DoctorPreferenceReport = doctor_preference_report,
    HETIComplianceTable = HETI_compliance_table,
    PreferenceMatrix = preference_matrix_df,
    UnassignedTerms = unassigned_df
  )
  
  # Export to .xlsx file
  writexl::write_xlsx(tables_list, path = filename)
}

# Call the function with all the necessary data frames and the filename
export_tables(
  assignment_table = assignment_table,
  unassigned_df = unassigned_df,
  doctor_preference_report = doctor_preference_report,
  HETI_compliance_table = HETI_compliance_table,
  preference_matrix = preference_matrix,
  evaluation_table = evaluation_table,
  
  filename = "PGY1_Allocations_export.xlsx"
)

```


*This code, inclusive of all associated documentation and related content, is the exclusive intellectual property of Dr. Scott Johnson Gelzinnis under the Copyright Act 1968 (Cth) of Australia. Any unauthorized use, reproduction, distribution, or modification of this material, in entirety or in parts, without the prior written consent of Dr. Scott Johnson Gelzinnis, may constitute a breach of Australian copyright laws and can result in legal action.*
