---
title: "PGY2 Solve - Allocations 2024"
author: "Dr Scott Johnson Gelzinnis"
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: journal
    df_print: paged
  html_document:
    theme: journal
    df_print: paged
  pdf_document: default
---

*Copyright Dr Scott Johnson Gelzinnis (2023) - Statement below*


  
# Setting up the Environment  


```{r include=TRUE, echo=TRUE, results='hide'}

library(here())
i_am("PGY2_Solve.Rmd")

#Set WD
# setwd("D:/My Documents/GitHub/term-allocations")

# List of required packages
required_packages <- c("randomNames", "lpSolve", "lpSolveAPI", "ROI", "gt", "readr", "tidyverse", "googlesheets4", "stringr", "knitr", "ompr", "ompr.roi", "ROI.plugin.glpk")

# Check if packages are installed and install them if they are not
new_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

# Load the packages
lapply(required_packages, require, character.only = TRUE)

# Random seed
set.seed(69)

```



# Loading in the Preference Matrix 

```{r include=TRUE, echo=TRUE, results='hide'}

## Read in Google Sheet
raw_preference_data <- read_sheet("https://docs.google.com/spreadsheets/d/1JW_KriVymSe9NNVi1QjKnshoWrkAhZ8FkKzpqXy1bf8/edit?usp=sharing")

# Convert the tibble to a data.frame
raw_preference_data <- as.data.frame(raw_preference_data)

## Define preference matrix  ##

# Combine columns 13 and 14 to create the "Doctor" column
raw_preference_data$Doctor <- paste(raw_preference_data[[13]], raw_preference_data[[14]], sep = ", ")

# Extract column names that have the pattern '| '
term_columns <- grep("\\| ", names(raw_preference_data), value = TRUE)

# Select the relevant columns
selected_data <- raw_preference_data[, term_columns]

# Set "Doctor" column as row names
rownames(selected_data) <- raw_preference_data$Doctor

# Clean the column names by removing everything before '| ' (i.e., pipe followed by a space)
colnames(selected_data) <- gsub(".*\\| ", "", colnames(selected_data))

# Convert selected data to a matrix and exclude the "Doctor" column
preference_matrix <- as.matrix(selected_data)

# Define dimensions based on preference_matrix
num_columns <- ncol(preference_matrix)
num_doctors <- nrow(preference_matrix)
num_terms <- num_columns  

# Assuming the number of terms is equal to the number of columns in the matrix
row_names <- rownames(selected_data)
column_names <- colnames(selected_data)

```


# Definitions

```{r include=TRUE, echo=TRUE, results='hide'}
               
                         ## Linear Optimization - Defining the problem ##

term_master_df <- read_csv("TRIAL PGY2 2024.csv")

# Directly infer the number of doctors and terms from the preference matrix
num_doctors <- nrow(preference_matrix)
num_terms <- ncol(preference_matrix)

# Define the objective function coefficients (preferences)
objective_coeffs <- as.vector(preference_matrix)

# Simplifying max doctor allocations
max_doctors_per_terms <- term_master_df %>%
  select(matches("PGY2 Term \\d+ Allocation")) %>%
  map_df(~ as.vector(.))

# Helper function for tidy conversions
convert_and_name <- function(column_name) {
  term_master_df %>%
    pull(column_name) %>%
    as.factor() %>%
    set_names(term_master_df$`Term Name`)
}

# Using the helper function to define the vectors
clinical_structure_term <- convert_and_name('Clinical Team Structure')
specialty_status_per_term <- convert_and_name('Specialty Status')
sub_specialty_status_per_term <- convert_and_name('Subspecialty Status')
rural_status_term <- convert_and_name('Rurality')
term_classification1 <- convert_and_name('Term Classification')
term_classification2 <- convert_and_name('Term Classification 2')

# Helper function to retrieve the correct 'max doctors' vector for a given assignment number
get_max_doctors_for_term <- function(assignment_num) {
  if(assignment_num >= 1 & assignment_num <= 5) {
    return(max_doctors_per_terms[[assignment_num]])
  } else {
    stop("Invalid assignment number!")
  }
}


# Define unique specialties and sub-specialties
unique_specialty_statuses <- unique(specialty_status_per_term)
unique_sub_specialties <- unique(sub_specialty_status_per_term)


```



# Linear Optimization Model with Minimized Dissatisfaction

### With allowable constraints to provide viable solution (i.e if unable to solve, roll back constraint one at a time)

```{r include=TRUE, echo=TRUE, results='hide'}

# Assuming the dissatisfaction values are set from the preference matrix
dissatisfaction_values <- as.vector(preference_matrix)

# Initialization of assigned_terms_matrix and other tracking matrices
assigned_terms_matrix <- matrix(0, nrow=num_doctors, ncol=num_terms)
assigned_specialties <- matrix(0, nrow=num_doctors, ncol=length(unique_specialty_statuses))
assigned_sub_specialties <- matrix(0, nrow=num_doctors, ncol=length(unique_sub_specialties))

# Running the assignment process 5 times
for(assignment_num in 1:5) {
  
# Start defining the model
model <- MIPModel() %>%
    add_variable(x[i, j], i = 1:num_doctors, j = 1:num_terms, type = "binary") %>%
    set_objective(sum_expr(dissatisfaction_values[(i - 1) * num_terms + j] * x[i, j], 
                       i = 1:num_doctors, 
                       j = 1:num_terms), 
              sense = "min")
 
# Constraints
# 1. One assignment per doctor
add_constraint(sum_expr(x[i, j](), j = 1:num_terms) == 1, i = 1:num_doctors) #%>%
 
# # 3. Max doctors per term constraint
# add_constraint(sum_expr(x[i, j](), i = 1:num_doctors) <= get_max_doctors_for_term(assignment_num)[j], j = 1:num_terms) %>%
#  
# # 4. Prevent reassignment to same term
# add_constraint(x[i, j]() == 0, i = 1:num_doctors, j = 1:num_terms, subset = assigned_terms_matrix[i,j] == 1)
  
# Once the model is defined, you can solve it
result <- solve_model(model, with_ROI(solver = "glpk", verbose = TRUE))

# Reshape the solution into a matrix, based on its length
assignments <- matrix(result$solution, nrow=num_doctors)

# Print the assignments for current iteration
cat(paste("\nAssignment round:", assignment_num, "\n"))

for (i in 1:num_doctors) {
  assigned_term <- which(assignments[i, ] == 1)
  
  # Using row_names and column_names to print the doctor's and term's name
  if (length(assigned_term) > 0) {
    cat(sprintf("%s is assigned to %s\n", row_names[i], column_names[assigned_term]))
  } else {
    cat(sprintf("%s has no assignment\n", row_names[i]))
  }
}

# Updating the assigned_terms_matrix and specialty tracking matrices after assignment
for (i in 1:num_doctors) {
  for (j in 1:num_terms) {
    if (assignments[i,j] == 1) {
      assigned_terms_matrix[i, j] <- 1
      assigned_specialties[i, which(unique_specialty_statuses == specialty_status_per_term[j])] <- assigned_specialties[i, which(unique_specialty_statuses == specialty_status_per_term[j])] + 1
      assigned_sub_specialties[i, which(unique_sub_specialties == sub_specialty_status_per_term[j])] <- assigned_sub_specialties[i, which(unique_sub_specialties == sub_specialty_status_per_term[j])] + 1
    }
  }
}
}



```



# PGY2 Unassigned terms (YET TO BE ADDED)
```{r}

```


# MODEL EVALUATION

```{r include=TRUE, echo=TRUE, results='markup'}

# How many Doctors with one Top 10 preference

# Convert the preference matrix into a top 10 boolean matrix
top_10_matrix <- t(apply(preference_matrix, 1, function(row) row %in% sort(row, decreasing = FALSE)[1:10]))

# For each doctor, determine if the assigned terms are in their top 10 preferences
assignment_top_10_matrix <- matrix(0, nrow = num_doctors, ncol = num_assignments)

for (assignment_num in 1:num_assignments) {
  assignment_column <- paste0("Assignment", assignment_num)
  assigned_terms <- assignment_table[, assignment_column]
  
  # Convert the assigned term names back to their column indices
  assigned_indices <- match(assigned_terms, colnames(selected_data))
  
  for (i in 1:num_doctors) {
    assignment_top_10_matrix[i, assignment_num] <- top_10_matrix[i, assigned_indices[i]]
  }
}

# Convert the preference matrix into a top 5 boolean matrix
top_5_matrix <- t(apply(preference_matrix, 1, function(row) row %in% sort(row, decreasing = FALSE)[1:5]))

# Create a function to compute metrics for a given top_x matrix
get_metrics <- function(matrix, assignment_table, num_assignments) {
  assignment_matrix <- matrix(0, nrow = num_doctors, ncol = num_assignments)
  
  for (assignment_num in 1:num_assignments) {
    assignment_column <- paste0("Assignment", assignment_num)
    assigned_terms <- assignment_table[, assignment_column]
    
    # Convert the assigned term names back to their column indices
    assigned_indices <- match(assigned_terms, colnames(selected_data))
    
    for (i in 1:num_doctors) {
      assignment_matrix[i, assignment_num] <- matrix[i, assigned_indices[i]]
    }
  }
  
  return(rowSums(assignment_matrix))
}

top_10_counts_per_doctor <- get_metrics(top_10_matrix, assignment_table, num_assignments)
top_5_counts_per_doctor <- get_metrics(top_5_matrix, assignment_table, num_assignments)

# Extract all assigned terms for each doctor across the assignments
assigned_terms_all <- as.vector(t(assignment_table[, 2:(num_assignments + 1)]))

# Convert the assigned term names back to their column indices
assigned_indices_all <- match(assigned_terms_all, colnames(selected_data))

# Retrieve preference values using matrix indexing
assigned_preference_values <- matrix(0, nrow = num_doctors, ncol = num_assignments)
for (assignment_num in 1:num_assignments) {
  assignment_column <- paste0("Assignment", assignment_num)
  assigned_terms <- assignment_table[, assignment_column]
  
  # Convert the assigned term names back to their column indices
  assigned_indices <- match(assigned_terms, colnames(selected_data))
  
  for (i in 1:num_doctors) {
    assigned_preference_values[i, assignment_num] <- preference_matrix[i, assigned_indices[i]]
  }
}

average_assigned_preference <- mean(as.vector(assigned_preference_values))

# Calculate the metrics of interest for each doctor
range_per_doctor <- apply(assigned_preference_values, 1, range)
max_per_doctor <- range_per_doctor[2,]
min_per_doctor <- range_per_doctor[1,]
sd_per_doctor <- apply(assigned_preference_values, 1, sd)

# Compute the average values across all doctors
average_max = mean(max_per_doctor)
average_min = mean(min_per_doctor)
average_sd = mean(sd_per_doctor)

# Create a data frame with the evaluation metrics
evaluation_data <- data.frame(
  Metric = c("At least 1 term in Top 10 (%)", 
             "At least 2 terms in Top 10 (%)", 
             "At least 3 terms in Top 10 (%)", 
             "At least 1 term in Top 5 (%)",
             "Average Preference of Assigned Terms",
             "Average Max of Assigned Preferences",
             "Average Min of Assigned Preferences",
             "Average SD of Assigned Preferences"),
  Value = c(sum(top_10_counts_per_doctor > 0) / num_doctors * 100, 
            sum(top_10_counts_per_doctor >= 2) / num_doctors * 100, 
            sum(top_10_counts_per_doctor >= 3) / num_doctors * 100, 
            sum(top_5_counts_per_doctor > 0) / num_doctors * 100,
            average_assigned_preference,
            average_max,
            average_min,
            average_sd)
)

# Round the values to 2 decimal places
evaluation_data$Value <- round(evaluation_data$Value, 2)

# Generate the table using gt
evaluation_table <- gt(evaluation_data) %>% 
  tab_header(
    title = "Model Evaluation"
  ) %>% 
  cols_label(
    Metric = "Metric",
    Value = "Value"
  )

# Print the table
evaluation_table


```


                         
# Reporting - 

### To add transparancy to allocations process and to assist in manual auditing of lines in order to preserve the integrity of the preferences of the cohort. Assist with manual changes to make terms HETI compliant or understand GAPS in the event their is nil viable solution to whole gamut of constraints. 

```{r include=TRUE, echo=TRUE, results='markup'}
# 1. Doctors Preference Report

# Initialize an empty data frame to store the report
doctor_preference_report <- data.frame(
  Doctor = character(),
  Term = character(),
  Preference_Score = numeric(),
  stringsAsFactors = FALSE
)

# Iterate through each doctor
for (i in 1:num_doctors) {
  doctor_name <- row_names[i]
  
  # Iterate through each of the assignments to extract the allocated terms for each doctor
  for (assignment_num in 1:num_assignments) {
    optimal_solution_dissatisfaction <- assignment_results[[assignment_num]]
    for (j in 1:num_terms) {
      if (optimal_solution_dissatisfaction[(i - 1) * num_terms + j] == 1) {
        assigned_term <- column_names[j]
        
        # Access the doctor's preference row from the preference matrix
        doctor_preference <- preference_matrix[i, ]
        
        # Calculate preference score for the assigned term
        preference_score <- doctor_preference[match(assigned_term, column_names)]
        
        # Build a data frame with the results for this doctor and term
        doctor_report <- data.frame(
          Doctor = rep(doctor_name, 1),
          Term = assigned_term,
          Preference_Score = preference_score
        )
        
        # Append the doctor's report to the overall report data frame
        doctor_preference_report <- rbind(doctor_preference_report, doctor_report)
      }
    }
  }
}


gt(doctor_preference_report)
```


```{r include=TRUE, echo=TRUE, results='markup'}
# 2. HETI Compliance Report


# Calculate average preference for each doctor across all 5 assignments
avg_preference <- doctor_preference_report %>%
  group_by(Doctor) %>%
  summarize(Avg_Score = mean(Preference_Score, na.rm = TRUE))

# Check if a doctor has been assigned to more than 1 term in any sub-specialty
sub_specialty_compliance <- sapply(1:num_doctors, function(i) {
  doctor_name <- row_names[i]
  assigned_terms_df <- subset(doctor_preference_report, Doctor == doctor_name)
  terms_assigned <- assigned_terms_df$Term
  sub_specialty_terms <- sub_specialty_status_per_term[terms_assigned]
  max_count <- max(table(sub_specialty_terms), na.rm = TRUE)
  return(max_count <= 1)
})

# Check if a doctor has been assigned to more than 2 terms in any specialty
specialty_compliance <- sapply(1:num_doctors, function(i) {
  doctor_name <- row_names[i]
  terms_assigned <- doctor_preference_report$Term[doctor_preference_report$Doctor == doctor_name]
  specialty_terms <- specialty_status_per_term[terms_assigned]
  max_count <- max(table(specialty_terms), na.rm = TRUE)
  return(max_count <= 2)
})

# Check if a doctor has been assigned to at least 3 "Team Based" terms
team_based_compliance <- sapply(1:num_doctors, function(i) {
  doctor_name <- row_names[i]
  terms_assigned <- doctor_preference_report$Term[doctor_preference_report$Doctor == doctor_name]
  
  # Debugging print statements
  print(paste("Doctor:", doctor_name))
  print(paste("Assigned terms:", toString(terms_assigned)))
  structure_indices <- match(terms_assigned, names(clinical_structure_term))
  print(paste("Matching indices:", toString(structure_indices)))
  print(paste("Clinical structures:", toString(clinical_structure_term[structure_indices])))
  
  team_based_count <- sum(clinical_structure_term[structure_indices] == "Team Based")
  
  return(team_based_count >= 3)
})

# Check for compliance with the desired term classifications
check_term_classifications <- function(terms_assigned) {
  classification1 <- term_classification1[match(terms_assigned, column_names)]
  classification2 <- term_classification2[match(terms_assigned, column_names)]
  required_terms <- c("A-Undifferentiated illness patient care", 
                      "B-Chronic illness patient care", 
                      "C-Acute critical illness patient care")
  
  # Check for each term individually, without merging the classifications
  has_required_terms <- sapply(required_terms, function(term) {
    term %in% classification1 || term %in% classification2
  })
  
  return(all(has_required_terms))
}

term_classifications_compliance <- sapply(1:num_doctors, function(i) {
  doctor_name <- row_names[i]
  terms_assigned <- doctor_preference_report$Term[doctor_preference_report$Doctor == doctor_name]
  return(check_term_classifications(terms_assigned))
})

compliance_data <- data.frame(
  Doctor = avg_preference$Doctor,
  Avg_Preference = avg_preference$Avg_Score,
  SubSpecialty_Compliance = sub_specialty_compliance,
  Specialty_Compliance = specialty_compliance,
  Team_Based_Compliance = team_based_compliance,
  Term_Classifications_Compliance = term_classifications_compliance
)

# Format with gt

HETI_compliance_table <- compliance_data %>%
  gt() %>%
  cols_label(
    Avg_Preference = "Avg. Preference (1-5)",
    SubSpecialty_Compliance = "Max 1 Term in Sub-Specialty",
    Specialty_Compliance = "Max 2 Terms in Specialty",
    Team_Based_Compliance = "At Least 3 'Team Based' Terms",
    Term_Classifications_Compliance = "Essential Classifications"
  )

print(HETI_compliance_table)
```



*This code, inclusive of all associated documentation and related content, is the exclusive intellectual property of Dr. Scott Johnson Gelzinnis under the Copyright Act 1968 (Cth) of Australia. Any unauthorized use, reproduction, distribution, or modification of this material, in entirety or in parts, without the prior written consent of Dr. Scott Johnson Gelzinnis, may constitute a breach of Australian copyright laws and can result in legal action.*
