---
title: "R Notebook"
output: html_notebook
---

# Set WD
setwd("C:/Users/Scotty/Desktop/Allocations Linear Optimisation")

# Install and load the necessary packages
```{r}
install.packages("randomNames")
install.packages("lpSolve")
install.packages("lpSolveAPI")
install.packages("ROI")
install.packages("knitr")
install.packages("gt")
```

#Load Libraries
```{r}
library(readr)
library(tidyverse)
library(skimr)
library(lpSolve)
library(randomNames)
library(lpSolveAPI)
library(ROI)
library(knitr)
library(gt)
```

#Random Seed
```{r}
set.seed(69)
```

#Define preference matrix
```{r}
# Set the number of rows and columns
num_rows <- 100
num_columns <- 45

# Generate random human-like row names
row_names <- randomNames(num_rows)

# Create an empty preference matrix with row names
preference_matrix <- matrix(NA, nrow = num_rows, ncol = num_columns,
                            dimnames = list(row_names, paste0("Term ", 1:num_columns)))

# Fill the preference matrix with unique random numbers
for (i in 1:num_rows) {
  preference_matrix[i, ] <- sample(1:num_columns)
  while (any(duplicated(preference_matrix[i, ]))) {
    preference_matrix[i, ] <- sample(1:num_columns)
  }
}

# Read column names from CSV file starting from A2
column_names <- read_csv("PGY2 Solve.csv", col_names = FALSE, skip = 1)
column_names <- column_names$X1

# Update column names of preference matrix
colnames(preference_matrix) <- column_names
```

# Linear Optimisation
```{r}
# Define the number of doctors and terms
num_doctors <- nrow(preference_matrix)
num_terms <- ncol(preference_matrix)

# Define the objective function coefficients (preferences)
objective_coeffs <- as.vector(preference_matrix)

# Create a vector of the maximum number of doctors allowed for each term
max_doctors_per_term <- read_csv("PGY2 Solve.csv", col_names = FALSE, skip = 1)
max_doctors_per_term <- as.vector(max_doctors_per_term$X3)
```

#Linear Optimization with Minimized Dissatisfaction
```{r}
# Calculate dissatisfaction values based on preferences (higher value indicates higher dissatisfaction)
dissatisfaction_values <- num_columns - preference_matrix + 1

# Convert dissatisfaction values to a vector for the objective function
objective_coeffs_dissatisfaction <- as.vector(dissatisfaction_values)

# Create an empty LP model for minimizing dissatisfaction
lp_model_dissatisfaction <- make.lp(0, num_doctors * num_terms)

# Set binary decision variables (0 or 1)
set.type(lp_model_dissatisfaction, columns = 1:(num_doctors * num_terms), type = "binary")

# Set the objective function coefficients for minimizing dissatisfaction
set.objfn(lp_model_dissatisfaction, objective_coeffs_dissatisfaction)

# Add constraints to ensure one assignment per doctor
for (i in 1:num_doctors) {
  constr_one_assignment <- rep(0, num_doctors * num_terms)
  constr_one_assignment[((i - 1) * num_terms + 1):(i * num_terms)] <- 1
  add.constraint(lp_model_dissatisfaction, constr_one_assignment, type = "=", rhs = 1)
}

# Add constraints to limit the maximum number of doctors allocated to each term
for (j in 1:num_terms) {
  constr_max_doctors_per_term <- rep(0, num_doctors * num_terms)
  constr_max_doctors_per_term[j + seq(0, (num_doctors - 1) * num_terms, by = num_terms)] <- 1
  add.constraint(lp_model_dissatisfaction, constr_max_doctors_per_term, type = "<=", rhs = max_doctors_per_term[j])
}
```

# Reporting intial solution
```{r}
# Print the LP model to check its settings (optional)
print(lp_model_dissatisfaction)

# Solve the linear programming problem to minimize dissatisfaction
lp_result_dissatisfaction <- solve(lp_model_dissatisfaction)

# Check if the LP problem was solved successfully
if (lp_result_dissatisfaction == 0) {
  # Get the optimal solution from the lp_model_dissatisfaction object
  optimal_solution_dissatisfaction <- get.variables(lp_model_dissatisfaction)
  
  # Store doctor-term assignments in a data frame
  optimal_assignments <- data.frame(
    Doctor = character(),
    Term = character(),
    stringsAsFactors = FALSE
  )
  
  # Analyze the optimal solution
  for (i in 1:num_doctors) {
    for (j in 1:num_terms) {
      if (optimal_solution_dissatisfaction[(i - 1) * num_terms + j] == 1) {
        doctor_name <- row_names[i]  # Get the doctor's name using the index
        term_name <- column_names[j]  # Get the term's name using the index
        optimal_assignments <- rbind(optimal_assignments, data.frame(Doctor = doctor_name, Term = term_name))
      }
    }
  }
  
  # Create a gt table to display the optimal solution
  optimal_assignments_table <- gt(optimal_assignments)
  print(optimal_assignments_table)
} else {
  cat("LP problem could not be solved. Result:", lp_result_dissatisfaction, "\n")
}
```

# NEXT STEP TO RUN A LOOP TO ASSIGN 5 TERMS ACROSS THE YEAR BASED ON PREFERENCES
```{r}
# Number of times to run the assignment
num_assignments <- 5

# Initialize a list to store the results of each assignment
assignment_results <- list()

# Loop through the assignments
for (assignment_num in 1:num_assignments) {
  # Create an empty LP model for minimizing dissatisfaction
  lp_model_dissatisfaction <- make.lp(0, num_doctors * num_terms)
  
  # Set binary decision variables (0 or 1)
  set.type(lp_model_dissatisfaction, columns = 1:(num_doctors * num_terms), type = "binary")
  
  # Set the objective function coefficients for minimizing dissatisfaction
  set.objfn(lp_model_dissatisfaction, objective_coeffs_dissatisfaction)
  
  # Add constraints to ensure one assignment per doctor
  for (i in 1:num_doctors) {
    constr_one_assignment <- rep(0, num_doctors * num_terms)
    constr_one_assignment[((i - 1) * num_terms + 1):(i * num_terms)] <- 1
    add.constraint(lp_model_dissatisfaction, constr_one_assignment, type = "=", rhs = 1)
  }
  
  # Add constraints to limit the maximum number of doctors allocated to each term
  for (j in 1:num_terms) {
    constr_max_doctors_per_term <- rep(0, num_doctors * num_terms)
    constr_max_doctors_per_term[j + seq(0, (num_doctors - 1) * num_terms, by = num_terms)] <- 1
    add.constraint(lp_model_dissatisfaction, constr_max_doctors_per_term, type = "<=", rhs = max_doctors_per_term[j])
  }
  
  # Solve the linear programming problem to minimize dissatisfaction
  lp_result_dissatisfaction <- solve(lp_model_dissatisfaction)
  
  # Check if the LP problem was solved successfully
  if (lp_result_dissatisfaction == 0) {
    optimal_solution_dissatisfaction <- get.variables(lp_model_dissatisfaction)
    assignment_results[[assignment_num]] <- optimal_solution_dissatisfaction
  } else {
    cat("LP problem could not be solved for assignment", assignment_num, "Result:", lp_result_dissatisfaction, "\n")
  }
}

# Print the results of each assignment
for (assignment_num in 1:num_assignments) {
  cat("Assignment", assignment_num, "results:\n")
  optimal_solution_dissatisfaction <- assignment_results[[assignment_num]]
  for (i in 1:num_doctors) {
    for (j in 1:num_terms) {
      if (optimal_solution_dissatisfaction[(i - 1) * num_terms + j] == 1) {
        doctor_name <- row_names[i]
        term_name <- column_names[j]
        cat("Doctor", doctor_name, "is assigned to Term", term_name, "\n")
      }
    }
  }
  cat("\n")
}

# Create a data frame to store the assignment results for each doctor
assignment_results_df <- data.frame(
  Doctor = row_names, 
  stringsAsFactors = FALSE
)

# Initialize a matrix to track assigned terms for each doctor
assigned_terms_matrix <- matrix(FALSE, nrow = num_doctors, ncol = num_terms)

# Iterate through the assignments
for (assignment_num in 1:num_assignments) {
  assignment_column <- paste0("Assignment ", assignment_num)
  assignment_results_df[[assignment_column]] <- NA
  
  for (i in 1:num_doctors) {
    # Filter unassigned terms for this doctor
    unassigned_terms <- column_names[!assigned_terms_matrix[i, ]]
    
    if (length(unassigned_terms) > 0) {
      # Calculate a combined score that considers preferences and dissatisfaction
      combined_score <- preference_matrix[i, ] - dissatisfaction_values
      
      # Sort terms based on the combined score
      sorted_terms <- unassigned_terms[order(combined_score[match(unassigned_terms, column_names)])]
      
      # Assign the doctor to the term with the highest combined score
      assigned_term <- sorted_terms[1]
      assigned_term_index <- match(assigned_term, column_names)
      
      # Mark the term as assigned for this doctor
      assigned_terms_matrix[i, assigned_term_index] <- TRUE
      
      # Update the data frame with the assigned term
      assignment_results_df[[assignment_column]][i] <- assigned_term
    }
  }
}

# Print the table using gt
assignment_results_table <- gt(assignment_results_df)
print(assignment_results_table)
                   